```{r}
library(tidyverse)
library(caret)
library(ggplot2)
rm(list = ls())#Clears environment variables
```
Reads in the .csv file which has already been added to the project folder
```{r}
ONP_Original <- read.csv(file = 'OnlineNewsPopularity.csv')
```

Removes Unnecessary Columns
```{r}
ONP_Trimmed <- ONP_Original%>%select(-2:-3,-5:-9,-12:-13,-20:-28,-40:-44)
```

Groups days of the week into one column for use in the models. The unite function introduced underscores and zeroes so they were removed using gsub. 
```{r}
ONP_Test <- ONP_Trimmed

ONP_Test$weekday_is_monday <- ONP_Test$weekday_is_monday%>%replace(ONP_Test$weekday_is_monday==1, "Monday")

ONP_Test$weekday_is_tuesday <- ONP_Test$weekday_is_tuesday%>%replace(ONP_Test$weekday_is_tuesday==1, "Tuesday")

ONP_Test$weekday_is_wednesday <- ONP_Test$weekday_is_wednesday%>%replace(ONP_Test$weekday_is_wednesday==1, "Wednesday")

ONP_Test$weekday_is_thursday <- ONP_Test$weekday_is_thursday%>%replace(ONP_Test$weekday_is_thursday==1, "Thursday")

ONP_Test$weekday_is_friday <- ONP_Test$weekday_is_friday%>%replace(ONP_Test$weekday_is_friday==1, "Friday")

ONP_Test$weekday_is_saturday <- ONP_Test$weekday_is_saturday%>%replace(ONP_Test$weekday_is_saturday==1, "Saturday")

ONP_Test$weekday_is_sunday <- ONP_Test$weekday_is_sunday%>%replace(ONP_Test$weekday_is_sunday==1, "Sunday")

ONP_Test <- ONP_Test%>%unite(weekday_is_monday:weekday_is_sunday, col = "day of week", remove = T, sep ="_" )

ONP_Test$`day of week` <- gsub("_", "",ONP_Test$`day of week`)
ONP_Test$`day of week` <- gsub("0", "", ONP_Test$`day of week`)
```

Groups data channels into one column for use with models. The unite function introduced underscores and zeroes so they were removed using gsub.
```{r}
ONP_Test$data_channel_is_lifestyle <- ONP_Test$data_channel_is_lifestyle%>%replace(ONP_Test$data_channel_is_lifestyle==1, "Lifestyle")

ONP_Test$data_channel_is_entertainment <- ONP_Test$data_channel_is_entertainment%>%replace(ONP_Test$data_channel_is_entertainment==1, "Entertainment")

ONP_Test$data_channel_is_bus <- ONP_Test$data_channel_is_bus%>%replace(ONP_Test$data_channel_is_bus==1, "Business")

ONP_Test$data_channel_is_socmed <- ONP_Test$data_channel_is_socmed%>%replace(ONP_Test$data_channel_is_socmed==1, "Social Media")

ONP_Test$data_channel_is_tech <- ONP_Test$data_channel_is_tech%>%replace(ONP_Test$data_channel_is_tech==1, "Technology")

ONP_Test$data_channel_is_world <- ONP_Test$data_channel_is_world%>%replace(ONP_Test$data_channel_is_world==1, "World")

ONP_Test <- ONP_Test%>%unite(data_channel_is_lifestyle:data_channel_is_world, col = "Article_Channel", remove = T, sep ="_" )

ONP_Test$Article_Channel <- gsub("_", "",ONP_Test$Article_Channel)
ONP_Test$Article_Channel <- gsub("0", "", ONP_Test$Article_Channel)
```

Removes rows with blanks in the Article Channel column,Check day of week column for blank rows then rename day of week to "Day_of_Week" 
```{r}
sum(ONP_Test$Article_Channel=="")#counts blanks in Article channel
ONP_Final <- ONP_Test[-which(ONP_Test$Article_Channel == ""), ]#removes rows with blanks
sum(ONP_Test$`day of week`=="")#all rows in Day of Week appear to be populated 
sum(ONP_Test$shares=="")
ONP_Final <- rename(ONP_Final, 'Day_of_Week' ='day of week')
sapply(ONP_Final, function(x) sum(is.na(x)))#checks for NA's
```
Creates a new column based on shares, called popularity. If the shares are above 2520.443 the article is considered popular. This column was created because the SVM model requires categorical attribute.
```{r}
ONP_Final$popularity = ifelse(ONP_Final$shares>2520.443,1,0)
ONP_Final$popularity <- as.factor(ONP_Final$popularity)
```

Sample of df specifically for running the SVM model, the original ~33k rows was taking extremely long to run. 
```{r}
ONP_Final <- ONP_Final[sample(nrow(ONP_Final), 10000), ]
```

Creates train and test data frames for the SVM Model.
```{r}
set.seed(111)
trainList <- createDataPartition(y=ONP_Final$popularity, p=.80,list=F)
trainSet <- ONP_Final[trainList,]
testSet <- ONP_Final[-trainList,]

svm.model <- train(popularity~Article_Channel+Day_of_Week+n_tokens_content+num_imgs+num_videos+rate_positive_words, data=trainSet, method="svmRadial", preProc=c("center", "scale"))
```

Prediction and confusion matrix for SVM model.
```{r}
svmPred <- predict(svm.model, newdata=testSet)
confusion1 <- confusionMatrix(svmPred, testSet$popularity)
confusion1
```

Linear Regression Model using seven attributes as predictors. 
```{r}
ONP.lm = lm(formula=shares ~ num_videos + num_imgs + n_tokens_content + rate_positive_words+rate_negative_words+Day_of_Week+Article_Channel, data = ONP_Final)

summary(ONP.lm)
```

Graph of day of week vs shares.
```{r}
positions <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday")
ggplot(ONP_Final) + geom_col(aes(x=(Day_of_Week),y=sum(shares)), color='palegreen4')+scale_x_discrete(limits = positions)
```

Graph videos vs shares.
```{r}
ggplot(ONP_Final) + geom_col(aes(x=num_videos, y=sum(shares)), color='palegreen4')+xlim(0,20)+ylim(0,1e+11)
```

Graph of images vs shares.
```{r}
ggplot(ONP_Final) + geom_col(aes(x=num_imgs, y=sum(shares)), color='palegreen4')+xlim(0, 30)
```

Article channel vs day of week.
```{r}
ggplot(ONP_Final) + geom_col(aes(x=Article_Channel, y=sum(shares)), color='palegreen4')
```



```{r}
predDF <- data.frame(num_videos=4, num_imgs=5, n_tokens_content=500, rate_positive_words=.8,rate_negative_words=.3,Day_of_Week='Monday',Article_Channel="Business")

predict(ONP.lm, predDF)
```



